---
title: "Climate Data ~ NFL"
author: "Katie Birchard"
date: '2018-04-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We have climate data for populations from the big dataset, containing all populations from Atlantic Canada, but now we want to take a look at climate data for just the NFL populations. 

* Since the lat and longs were not available in the paper (Bradbury et al. 2015), Kurtis put together a CSV file containing all the lats and longs for all Newfoundland populations - from both datasets used

```{r}
# Load dataset with all the lats and longs
NFLclim <- read.csv("NFLD rivers lat longCSV.csv")
View(NFLclim)
NFLclim <- NFLclim[,apply(NFLclim, 2, function(x) { sum(!is.na(x)) > 0 })] # function to remove columns that just contain NA
View(NFLclim)

# need to load raster to get access to worldclim data
library(raster)

rast <- getData("worldclim", var="bio", res=10) #making raster object, using biological variables, lowest resolution
rast <- rast[[c(1)]] #taking out only the mean annual temperature variable (taken b/w 1950-2000)

coords <- data.frame(x=NFLclim$Longitude, y=NFLclim$Latitude) #making dataframe for my lats and longs

points <- SpatialPoints(coords, proj4string = rast@crs) #determining the datapoints by specifying my lat and long

values <- extract(rast, points) #extracting the mean annual temperature for those lats and longs

df <- cbind.data.frame(coordinates(points), values) #putting lat, long, and mean annual temp into a dataframe

head(df) #this dataframe includes temperatures for all specific lats and longs for all populations

df$values <- values/10 #temp is in degrees C*10, so dividing by 10 to get actual degrees C
head(df)

projection(rast)=CRS("+init=epsg:4326") #want to make sure I'm using the right coordinate system

quartz()
plot.new()
plot(rast$bio1, xlim=c(-80,-50), ylim=c(45,58), 1, main="", xaxt="n", yaxt="n", legend=F) #plotting the newfoundland populations to get an idea of where they are  
points(cbind(NFLclim$Longitude, NFLclim$Latitude), col="#003F91FF", cex=1.5, pch=16) #adding the points
```

Now I'm going to make the dataframe including lat, long, and temperature values for current, past, and future temperatures:

```{r}
#df is the dataframe containing current temperature values
CurTemp <- df$values

#Now to get future temperature data, the same as for the big file:
rast2 <- getData("CMIP5", var="bio", res=10, rcp=85, model="GS", year=50) #used rcp=85 based on Sanford et al. 2014, used model=GISS based on Miao et al. 2014
rast2 <- rast2[[c(1)]]
names(rast2) <- "futTemp"
values2 <- extract(rast2, points)
View(values2)

FutTemp <- values2/10# temperature is in degrees Celsius *10, so need to divide by 10
head(FutTemp)

#Now to get past climate data:
pastClim <- read.table("pastData.txt", header=T) #Using GISS model to get past climate data from pre-industrial period (1770)
#whoops I accidentally deleted pastData.txt from the github - restored this using terminal
head(pastClim)

NFLclim$pastTemp <- NA #making a new column in the dataset for the past data
NFLclim_rounded <- NFLclim
NFLclim_rounded$Latitude <- ceiling(NFLclim_rounded$Latitude)-0.25  
NFLclim_rounded$Longitude <- ceiling(NFLclim_rounded$Longitude) + 0.25 #have to round to the .75 so I can match it with the PastClim data
View(NFLclim_rounded) #now ready to put into loop

#loop will take forever to run because PastClim has >250,000 rows, so I will cut out unnecessary rows (rows outside range)
range(pastClim$lat)
range(NFLclim_rounded$Latitude) # only need pastClim$lat between 46.75 and 55.75

pastClim2 <- pastClim[which(pastClim$lat >= 46.75 & pastClim$lat <= 55.75), ]
nrow(pastClim2) # much better

range(pastClim2$long)
range(NFLclim_rounded$Longitude) #only need pastClim$long between -60.75 and -52.75

pastClim3 <- pastClim2[which(pastClim2$long >= -60.75 & pastClim2$long <= -52.75), ]
nrow(pastClim3) #323 rows, so should not take as long to cycle through the loop

for (i in 1:nrow(pastClim3)) { #making a loop to populate the pastTemp column with the correct temperatures
	for(j in 1:nrow(NFLclim_rounded)) {
		if (NFLclim_rounded[j,]$Latitude == pastClim3[i,]$lat & NFLclim_rounded[j,]$Longitude == pastClim3[i,]$long) {
			NFLclim_rounded[j, ]$pastTemp <- pastClim3[i,]$bio.1	
	}
	
	}
}

# Now add all the climate data together!

NFLclim_rounded$FutTemp <- FutTemp
NFLclim_rounded$CurTemp <- CurTemp

#now turning it into a csv

write.csv(NFLclim_rounded, file="ClimTrendDataNFL.csv")
```


