---
title: "Climate Regressions ~ NFL"
author: "Katie Birchard"
date: '2018-04-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

That was with the big data file containing all the populations across Atlantic Canada - now I'm going to do the same thing with just the newfoundland populations:

```{r}
NFLclim <- read.csv("ClimTrendDataNFL.csv")
View(NFLclim) # must have forgotten to run a line of code in the other code, Curtemp should only be for bio.1
#quick fix:
NFLclim <- subset(NFLclim, select= -c(CurTemp.bio2:CurTemp.bio19))
colnames(NFLclim)[colnames(NFLclim)=="CurTemp.bio1"] <- "CurTemp"
View(NFLclim)

NFLcoords <- read.csv("NFLD rivers lat longCSV.csv", header =T) # Kurtis' dataset with all the lats and longs
View(NFLcoords) # needs to be fixed - there are some duplicates of rivers in here and some of the lats and longs are off

# Simms Brook Fortune Bay
NFLcoords$Longitude[18] <- -55.48
#Old Bay Brook
NFLcoords$Longitude[19] <- -55.60
# Connie River
NFLcoords$Longitude[20] <- -55.70
#Connie River Mill Brook
NFLcoords$Longitude[21] <- -55.65
# Little River
NFLcoords$Longitude[22] <- -55.81
# Dolland Brook
NFLcoords$Longitude[26] <- -57.75
# Garndy's Bay of Islands
NFLcoords$Longitude[27] <- -58.85
# La Poile River 
NFLcoords$Longitude[31] <- -58.32

# Getting rid of duplicate rivers
NFLclimFixed <- NFLclim[-c(32, 23, 49, 16, 29, 13), ]
View(NFLclimFixed)
NFLcoordsFixed <- NFLcoords[-c(32, 23, 49, 16, 29, 13), ]
View(NFLcoordsFixed)

NFLtotal <- merge(NFLclimFixed, NFLcoordsFixed, by="River")
View(NFLtotal)
NFLtotal2 <- NFLtotal[grep("^X", colnames(NFLtotal), invert = TRUE)]
head(NFLtotal2)
colnames(NFLtotal2)[which(names(NFLtotal2) == "Latitude.x")] <- "Latitude_rounded"
colnames(NFLtotal2)[which(names(NFLtotal2) == "Latitude.y")] <- "Latitude"
colnames(NFLtotal2)[which(names(NFLtotal2) == "Longitude.x")] <- "Longitude_rounded"
colnames(NFLtotal2)[which(names(NFLtotal2) == "Longitude.y")] <- "Longitude"
View(NFLtotal2)

#we also need river codes 

write.csv(NFLtotal2, file="ClimTrendDataNFL3.csv")

# afterwards, realized that these climate trends will be wrong for these locations...have to redo that part in new file (see ClimDataNFL2.Rmd)

# NFLData <- read.csv("ClimTrendDataNFL3.csv") --> updated file, now use version 4
NFLData <- read.csv("ClimTrendDataNFL4.csv")
head(NFLData)
NFLData1 <- NFLData[grep("^X", colnames(NFLData), invert = TRUE)] #not sure why these X rows keep appearing...

NFLData2 <- NFLData1[order(NFLData1$River), ]
head(NFLData2)
View(NFLData2)

NFLData2$Code <- c("NL-ALX", "NL-BBR", "BDN", "NL-BVB", "NL-BIG", "BBA", "NL-BER", "BSB", "NL-BDB", "NL-CMP", "CCR", "CNR", "CMB", "NL-DWB", "DHB", "NL-ENG", "NL-RBR", "NL-FLB", "GBI", "GAR", "GBB", "NL-COD", "GNR", "NL-GRB", "GRR", "NL-HUR", "NL-IAB", "NL-IBB", "IAM", "NL-JUB", "NL-KEN", "LPR", "NL-LJB", "LRD", "LSR", "NL-LOM", "LHR", "NL-MNR", "NL-MBA", "NHR", "NL-NBT", "NPR", "NWT", "OBB", "NL-PGB", "NL-PIN", "NL-PHR", "NL-RWR", "NL-ROP", "RKR", "NL-ROF", "NL-SRN", "SLR", "NL-SAN", "FBS", "NL-GSB", "SPR", "SWB", "SBM", "NL-SGR", "NL-SMR", "STS", "STN", "NL-TYB", "NL-TNR", "NL-WAB", "WBR")
View(NFLData2)

NFLData3 <- NFLData2[order(NFLData2$Code), ]
View(NFLData3)

HWEvalues <- read.csv("HWEdatNFL.csv") #downloading the HWE dataset that contains values for allelic richness, expected heterozygosity, and HWE p-values
View(HWEvalues)

#need to take out row 3 and 24, as these populations are not in the other dataset
HWEvalues2 <- HWEvalues[-c(3, 24), ]

HWEvalues3 <- HWEvalues2[order(HWEvalues2$Populations), ] #reordering the populations by alphabetizing them so they line up with the climate dataset

HWEvalues3$pastTemp <- NFLData3$pastTemp #adding past temp data
HWEvalues3$CurTemp <- NFLData3$CurTemp #adding current temp data
HWEvalues3$FutTemp <- NFLData3$FutTemp #adding future temp data

alldat <- HWEvalues3 #renaming the dataset to something more appropriate
View(alldat) #making sure it's in the correct order 

#need climate trends as well (i.e. slopes of climate data over time)
trenddat <- data.frame(1:nrow(alldat), alldat$pastTemp, alldat$CurTemp, alldat$FutTemp)

#making a function for slope
slope  <-  function(x){
    if(all(is.na(x)))
        return(NA) #returns NA if missing data
    else
        return(coef(lm(I(1:3)~x))[2]) #uses a linear model to find the slope of 3 consecutive values (Past temp, Cur temp, and fut temp)
}

# now apply the function to all rows in trenddat and insert the slope into a new column called "trends"
trenddat$trends <- apply(trenddat[ ,c("alldat.pastTemp", "alldat.CurTemp", "alldat.FutTemp")], 1, slope)

alldat$ClimTrend <- trenddat$trends #adding the trend column to the big dataset

head(alldat) #yay, it worked!
View(alldat)
```

Now, we will see if there is a correlation between climate trends and allelic richness.

* We want the climate trends to be our predictor variable and allelic richness to be our outcome variable

* Basically trying to see if allelic richness can be "predicted" by changes in mean annual temperature

```{r}
# first, we will do a quick visualization with a scatterplot to see if there are any apparent trends:
scatter.smooth(x=alldat$ClimTrend, y=alldat$Allelic.Richness, main="Allelic Richness vs. Temp Trends")
# seems like there is a bit more variation with negative climate trend slopes

# Now I will quickly check for outliers in the data using a boxplot:
boxplot(alldat$ClimTrend, main="Temp Trends", sub=paste("Outlier rows: ", boxplot.stats(alldat$ClimTrend)$out))
#ClimTrend doesn't seem to have any outliers
boxplot(alldat$Allelic.Richness, main="Allelic Richness", sub=paste("Outlier rows: ", boxplot.stats(alldat$Allelic.Richness)$out))
#Allelic Richness has a few outliers...

boxplot.stats(alldat$Allelic.Richness)$out #7.87 and 8.25 are all listed as outliers

#Now let's check for normality by doing a density plot
# install.packages("e1071") #already installed package in ClimateRegressions.Rmd
library(e1071)
plot(density(alldat$ClimTrend), main= "Density Plot: Temp Trends", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$ClimTrend), 2)))
# the plot for temp trends is slightly skewed, but there are no outliers so it should be ok...
# skewedness = 0.73

plot(density(alldat$Allelic.Richness), main= "Density Plot: Temp Trends", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$Allelic.Richness), 2)))
# Skewedness = -1.22

#get a look at the histogram
hist(alldat$Allelic.Richness)

# Now lets look at the level of correlation between Allelic Richness and Temp Trends
## first with the outliers
cor(alldat$ClimTrend, alldat$Allelic.Richness) # R = 0.02723

## now without the outliers
ARoutliers <- boxplot.stats(alldat$Allelic.Richness)$out
min(ARoutliers)
max(ARoutliers)

ARdrop <- alldat[!(alldat$Allelic.Richness <= max(ARoutliers)),]
cor(ARdrop$ClimTrend, ARdrop$Allelic.Richness) # R= -0.0666
# there's a bit more of a correlation, but still not much

#how does this hist look:
hist(ARdrop$Allelic.Richness)
shapiro.test(ARdrop$Allelic.Richness) #data barely fails normality test! (p=0.04853)
shapiro.test(log10(ARdrop$Allelic.Richness)) #data not normal (p=0.005357)
shapiro.test(1/(ARdrop$Allelic.Richness)) #not normal (p=0.000371) 
# No transformations necessary, shapiro test should be used more as a guideline anyway than a rule

# lets see what the scatterplot looks like for this:
scatter.smooth(x=ARdrop$ClimTrend, y=ARdrop$Allelic.Richness, main="Allelic Richness vs. Temp Trends")
# yeahhhhhh, still not really looking like a correlation

# Now I'm making a linear model to get a better quanititative idea
lmAR <- lm(Allelic.Richness ~ ClimTrend, data=alldat)
print(lmAR) # gives intercept and slope
summary(lmAR) # gives summary statistics, p=0.827 (not significant), R^2=0.0007412 ... no correlation
resAR <- residuals(lmAR) #finding residuals
plot(alldat$ClimTrend, resAR) #plotting residuals(difference b/w observed value and predicted value)

#what about a linear model for the dataset with the outliers removed:
lmARdrop <- lm(Allelic.Richness ~ ClimTrend, data=ARdrop)
summary(lmARdrop) #p=0.598, R^2=0.0004439 ... no correlation still

#I want to make a ggplot figure to show the linear regression (or lack of)
library(ggplot2)

#customizing a ggplot theme
install.packages("extrafont") # I want to use a different font from the base font (Helvetica)
library(extrafont)
font_import("Arial") # didn't work

install.packages("showtext")
library(showtext)
font_add_google("Arial")

theme_katie<-function(base_size = 18, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace% 
    theme(axis.text = element_text(size = rel(0.8),colour="black"),
          axis.ticks = element_line(colour = "black"),
    			legend.key = element_rect(fill="transparent", colour=NA),
    			panel.background = element_blank(),
    			legend.background = element_rect(fill="transparent", colour=NA),
          legend.position = "none",
          panel.border = element_rect(fill = NA, colour = NA), 
          panel.grid.major = element_line(colour = NA,size = 0), 
          panel.grid.minor = element_line(colour = NA,size = 0), 
          axis.line = element_line(colour ="black")
    )
}

install.packages("ggpmisc") #installing package that will allow me to put the p-values and R^2 values directly on the plots
library(ggpmisc)

lm_eqn <- function(df, y, x){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(R)^2~"="~R2, 
         list(a = format(coef(m)[1], digits = 2), 
              b = format(coef(m)[2], digits = 2), 
             R2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}

plotlmARdrop <- lm_eqn(ARdrop, ARdrop$Allelic.Richness, ARdrop$ClimTrend)
plotlmARdrop

ARbigLM <- ggplot(ARdrop, aes(y=Allelic.Richness, x=ClimTrend)) +
	geom_point(colour="cyan3", size=5, alpha=0.6) +
	#geom_smooth(method=lm, colour="cyan4") +
	theme_katie() +
	ylab("Allelic Richness") +
	xlab("Trend in Annual Temperature over Time") +
	stat_fit_glance(method="lm", geom="text", aes(label=paste0("p = ", round(..p.value.., 3))), label.x.npc = 0.8, label.y.npc = 0.2) +
	stat_poly_eq(formula = ARdrop$Allelic.Richness ~ ARdrop$ClimTrend, 
                aes(label = paste0(..rr.label..)), label.x.npc = 0.8, label.y.npc = 0.1, parse = TRUE)
# ARbigLM + geom_text(x = 0.6, y = 14, label = lm_eqn(ARdrop, ARdrop$Allelic.Richness, ARdrop$ClimTrend), parse = TRUE) --> this doesn't really match the style
print(ARbigLM)
ggsave("AR_NFL.pdf")
```


```{r}
#what if we log transform it?
lmARlog <- lm(log10(Allelic.Richness) ~ log10(ClimTrend), data=alldat)
print(lmARlog)
summary(lmARlog) #p=0.6708

#boostrap simulation b/w populations with neg climate trend vs. pos clim trend?
## trying to find out if mean allelic richness is significantly different between pos and neg climate trends

# make new column in alldat for categorical variable - clim pos/neg
alldat$ClimSlope <- NA
head(alldat)

#Making a loop to assign either a 0 or a 1 to the populations for my new column
for (i in 1:nrow(alldat)) {
	if (alldat$ClimTrend[i] > 0) {
		alldat$ClimSlope[i] <- "Pos"
	} 
	if (alldat$ClimTrend[i] < 0) {
		alldat$ClimSlope[i] <- "Neg"
	} 
}

# Clim slope is a binary variable so I can do a quick logistic regression to see if there might be an association between slope and allelic richness
logres1 <- glm(ClimSlope~Allelic.Richness, data=alldat, family=binomial(link="logit"))
summary(logres1)

Iters <- 1000 #we want to take a random sample 1000 times
BootOut <- rep(NA, Iters) #making a column with 1000 levels for the output of my bootstrap analysis
ClimPos <- alldat$Allelic.Richness[grep("Pos", alldat$ClimSlope)] #taking all the allelic richness values that have positive climate slope values
ClimNeg <- alldat$Allelic.Richness[grep("Neg", alldat$ClimSlope)] #taking all the allelic richness values that have a negative slope
for (i in 1:Iters){
	BootOut[i] <-
		mean(sample(ClimPos, length(ClimPos), replace=T)) - #taking mean of a sample of the allelic richness values for positive slopes
		mean(sample(ClimNeg, length(ClimNeg), replace=T)) #subtracting the mean of a sample of allelic richness values for negative slopes
}
head(BootOut) 


qplot(BootOut) # look at the distribution of values

XBootOut <- mean(BootOut) #mean of distribution is 0.0149
CIlow <- sort(BootOut)[floor(length(BootOut)*0.025)] 
CIhigh <- sort(BootOut)[ceiling(length(BootOut)*0.975)]
CIboot <- round(c(CIlow, CIhigh),2) #[-0.55, 0.59]

#To find p-value, need the mean difference of the raw data:
RawDat <- aggregate(alldat$Allelic.Richness, list(alldat$ClimSlope), mean)
RawDat

RawDiff <- RawDat$x[2] - RawDat$x[1]
qplot(BootOut) + theme_bw() + geom_vline(aes(xintercept=RawDiff), colour="blue")

p <- sum(BootOut>RawDiff)/length(BootOut)
paste("The estimated p-value is:", p) #p=0.489, so not really a big difference here

#do bootstrap simulation b/w populations that are endangered and not endangered
StatDat <- read.csv("EndangermentStatusNFL2.csv", header=T) #loading the csv file with the endangerment statuses
# we can do a logistic regression to see if allelic richness is related to endangerment status
head(alldat)
head(StatDat)
StatDat2 <- StatDat[order(StatDat$Code), ]
head(StatDat2)
View(StatDat2)
View(alldat) # now in the same order

alldat$Status <- StatDat2$Status

mod1 <- glm(Status~Allelic.Richness, data=alldat, family=binomial(link="logit"))
summary(mod1) #p=0.671, not significant at all

#adding status to ARdrop too
ARdrop <- alldat[!(alldat$Allelic.Richness <= max(ARoutliers)),]
#try without outliers
mod2 <- glm(Status~Allelic.Richness, data=ARdrop, family=binomial(link="logit"))
summary(mod2) #p=0.623

#see if there is a correlation with latitude/longitude
mod3 <- glm(Status~Latitude*Longitude, data=StatDat2, family=binomial(link="logit"))
summary(mod3) #significant coorelations here

#Maybe there is a correlation b/w allelic richness and latitude
alldat$Latitude <- StatDat2$Latitude
alldat$Longitude <- StatDat2$Longitude

lmARcoords <- lm(Allelic.Richness~Latitude*Longitude, data=alldat)
summary(lmARcoords) #not significant, p=0.5529

#adding latitude/longitude to ARdrop
ARdrop <- alldat[!(alldat$Allelic.Richness <= max(ARoutliers)),]

lmARcoords2 <- lm(Allelic.Richness~Latitude*Longitude, data=ARdrop)
summary(lmARcoords2) #once you account for outliers, still no significant correlation (p=0.48)

```
Next, we will see if there is a correlation between climate trends and expected heterozygosity in the Newfoundland populations:

```{r}
# first, we will do a quick visualization with a scatterplot to see if there are any apparent trends:
scatter.smooth(x=alldat$ClimTrend, y=alldat$Expected.Het, main="Expected Heterozygosity vs. Temp Trends")
# looks like expected heterozygosity is pretty stable with temperature fluctuations...

# Now I will quickly check for outliers in the data using a boxplot:
boxplot(alldat$Expected.Het, main="Expected Heterozygosity", sub=paste("Outlier rows: ", boxplot.stats(alldat$Expected.Het)$out))
#There are a few outliers for expected heterozygosity

boxplot.stats(alldat$Expected.Het)$out #0.70 0.77 0.75 are all listed as outliers

#Now let's check for normality by doing a density plot
plot(density(alldat$Expected.Het), main= "Density Plot: Expected Heterozygosity", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$Expected.Het), 2)))
# this is even more skewed than the allelic richness density plot
# Skewedness = -2.4

# Now lets look at the level of correlation between Expected Heterozygosity and Temp Trends
## first with the outliers
cor(alldat$ClimTrend, alldat$Expected.Het) # R = 0.103

## now without the outliers
HEoutliers <- boxplot.stats(alldat$Expected.Het)$out
min(HEoutliers)
max(HEoutliers)

HEdrop <- alldat[!(alldat$Expected.Het <= max(HEoutliers)),]
cor(HEdrop$ClimTrend, HEdrop$Expected.Het) # R= -0.018

# lets see what the scatterplot looks like for this:
scatter.smooth(x=HEdrop$ClimTrend, y=HEdrop$Expected.Het, main="Expected Heterozygosity vs. Temp Trends")
# no correlation here

# Now I'm making a linear model to get a better quanititative idea
lmHE <- lm(Expected.Het ~ ClimTrend, data=alldat)
summary(lmHE) # gives summary statistics, p=0.408, R^2=0.01056

#what if we log transform it?
lmHElog <- lm(log10(Expected.Het) ~ log10(ClimTrend), data=alldat)
summary(lmHElog) #p=0.9855, R^2<0.001

#linear model without the outliers:
lmHEdrop <- lm(Expected.Het ~ ClimTrend, data=HEdrop)
summary(lmHEdrop) #p=0.8872, R^2=0.000327

mod4 <- glm(Status~Expected.Het, data=alldat, family=binomial(link="logit"))
summary(mod4) #p=0.921, not significant

HEdrop <- alldat[!(alldat$Expected.Het <= max(HEoutliers)),]

mod5 <- glm(Status~Expected.Het, data=HEdrop, family=binomial(link="logit"))
summary(mod5) #p=0.390, still not significant

HEbigLM <- ggplot(HEdrop, aes(x=ClimTrend, y=Expected.Het)) +
	geom_point(colour="cyan3", size=5, alpha=0.6) +
	#geom_smooth(method=lm, colour="cyan4") +
	theme_katie() +
	ylab("Expected Heterozygosity") +
	xlab("Trend in Annual Temperature over Time") +
	stat_fit_glance(method="lm", geom="text", aes(label=paste0("p = ", round(..p.value.., 3))), label.x.npc = 0.8, label.y.npc = 0.2) +
	stat_poly_eq(formula = HEdrop$Expected.Het ~ HEdrop$ClimTrend, 
                aes(label = paste0(..rr.label..)), label.x.npc = 0.8, label.y.npc = 0.1, parse = TRUE)
print(HEbigLM)
ggsave("HE_NFL.pdf")

```

Then, we will see if there is a correlation between climate trends and deviations from HWE:

```{r}
# first, we will do a quick visualization with a scatterplot to see if there are any apparent trends:
scatter.smooth(x=alldat$ClimTrend, y=alldat$HWE.values, main="Deviations from HWE vs. Temp Trends")
# looks like there is no definitive relationship between temp trends and deviations from HWE...

# Now I will quickly check for outliers in the data using a boxplot:
boxplot(alldat$HWE.values, main="Deviations from HWE", sub=paste("Outlier rows: ", boxplot.stats(alldat$HWE.values)$out))
#no outliers...

range(alldat$HWE.values)
hist(alldat$HWE.values) #looks like there are 2 groups of individuals, one that deviates from HWE, and one that does not - should do logistical regression for this

# I'm going to make a nice histogram for the HWE p-values:
HWE_hist <- ggplot(data=alldat, aes(HWE.values)) + 
	geom_histogram(breaks=seq(0, 1, by=0.05),
								 col="cyan4", #fill="cyan3"
								 aes(fill=..count..),
								 alpha=0.8) +
  scale_fill_gradient("Populations", low = "cyan1", high = "cyan3") +
	theme_katie() +
	labs(title = "Newfoundland") +
	labs(x= "Hardy Weinberg Equilibrium P-values", y="Number of Populations") #+
	#geom_density(col="cyan4")
print(HWE_hist)
ggsave("HWEhist_NFL.pdf")

```

```{r}

#Now let's check for normality by doing a density plot
plot(density(alldat$HWE.values), main= "Density Plot: Deviations from HWE", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$HWE.values), 2)))
# bimodal distribution

# I can make the HWE p values truly dichotomous by assigning p < 0.05 == 0, and p > 0.05 == 1 - then I can run a logistic regression on it to see if there is a correlation with temperature trends

#I'm making a new row to put the binary HWE values in:
alldat$HWE.values2 <- NA
head(alldat)

#Making a loop to assign either a 0 or a 1 to the populations for my new column
for (i in 1:nrow(alldat)) {
	if (alldat$HWE.values[i] >= 0.05) {
		alldat$HWE.values2[i] <- 1
	} else {
		alldat$HWE.values2[i] <- 0
	} 
}

head(alldat) #my new column is populated, so now I can perform a logistic regression

mod1<-glm(HWE.values2~ClimTrend, data=alldat, family=binomial(link="logit"))
summary(mod1) # p=0.3516, not significant
# endangered populations tend to be further south, so maybe we should try with latitude

# Running a logistic regression with HWE deviations and latitude
mod2<-glm(HWE.values2~Latitude, data=alldat, family=binomial(link="logit"))
summary(mod2) #p=0.180

#Latitude + longitude
mod3 <- glm(HWE.values2~Latitude+Longitude, data=alldat, family=binomial(link="logit"))
summary(mod3) #not significant

#just longitude
mod4 <- glm(HWE.values2~Longitude, data=alldat, family=binomial(link="logit"))
summary(mod4) #0.238

# What about if there is a relationship b/w deviations from HWE and endangerment status?
M <- as.table(rbind(alldat$Status, alldat$HWE.values2))
dimnames(M) <- list(value = c("Status", "HW p-values"),
										population = alldat$Populations)
View(M)

chi1 <- chisq.test(M)
chi1 #p-value is 1

chi2 <- chisq.test(alldat$Status, alldat$HWE.values2, correct=F) #p=0.853
summary(chi2) #df =1, whereas df should be 66

# Also do a bootstrap similuation to determine differences between means for endangered and non-endangered

Iters <- 1000 #we want to take a random sample 1000 times
BootOut <- rep(NA, Iters) #making a column with 1000 levels for the output of my bootstrap analysis
HWDev <- alldat$HWE.values2[grep("Threatened", alldat$Status)] #taking all the Hardy Weinberg values for the threatened populations
HWFit <- alldat$HWE.values2[grep("Least Concern", alldat$Status)] #taking all the Hardy Weinberg values for the non-threatened populations

for (i in 1:Iters){
	BootOut[i] <-
		mean(sample(HWDev, length(HWDev), replace=T)) - #taking mean of a sample of HW values for threatened pops
		mean(sample(HWFit, length(HWFit), replace=T)) #subtracting the mean of a sample of HW values for non-threatened pops
}
head(BootOut) 


qplot(BootOut) # look at the distribution of values

XBootOut <- mean(BootOut) #mean of distribution is -0.0242
CIlow <- sort(BootOut)[floor(length(BootOut)*0.025)] #taking lower confidence interval
CIhigh <- sort(BootOut)[ceiling(length(BootOut)*0.975)]
CIboot <- round(c(CIlow, CIhigh),2) #-0.26, 0.20

#To find p-value, need the mean difference of the raw data:
RawDat <- aggregate(alldat$HWE.values2, list(alldat$Status), mean)
RawDat

RawDiff <- RawDat$x[2] - RawDat$x[1]
qplot(BootOut) + theme_bw() + geom_vline(aes(xintercept=RawDiff), colour="blue")

p <- sum(BootOut>RawDiff)/length(BootOut)
paste("The estimated p-value is:", p) #0.489, not significant, 
#use logistic regression

#plot on map
```
Interestingly, the Hardy-Weinberg p-values appear to be bimodally distributed, with populations either being significantly differentiated from Hardy-Weinberg equilibrium or not significantly differentiated from HWE.

* I'm going to plot this on a map showing the annual mean temperature

```{r}
library(raster)

rast <- getData("worldclim", var="bio", res=2.5) #making raster object, using biological variables, lowest resolution
rast <- rast[[c(1)]] #taking out only the mean annual temperature variable (taken b/w 1950-2000)

alldat$Latitude <- StatDat2$Latitude
alldat$Longitude <- StatDat2$Longitude

coords <- data.frame(x=alldat$Longitude, y=alldat$Latitude) #making dataframe for my lats and longs

points <- SpatialPoints(coords, proj4string = rast@crs) #determining the datapoints by specifying my lat and long

values <- extract(rast, points) #extracting the mean annual temperature for those lats and longs

df <- cbind.data.frame(coordinates(points), values) #putting lat, long, and mean annual temp into a dataframe

head(df) #this dataframe includes temperatures for all specific lats and longs for all populations

projection(rast)=CRS("+init=epsg:4326") #want to make sure I'm using the right coordinate system

quartz()
plot.new()

#want specific colour palette:
col.list <- c("cyan4", "deeppink1")
palette(col.list)

View(alldat)

threatened <- alldat[which (alldat$Status == "Threatened"), ]
nrow(threatened)

leastcon <- alldat[which (alldat$Status == "Least Concern"), ]
nrow(leastcon)

rastNFL <- plot(rast$bio1, xlim=c(-65,-50), ylim=c(45,58), 1, main="Newfoundland Populations", xaxt="n", yaxt="n", legend=F, asp=1) #plotting the newfoundland populations to get an idea of where they are 
rastNFL

points(cbind(threatened$Longitude, threatened$Latitude), col="black", cex=1, pch=4) #adding the points
points(cbind(leastcon$Longitude, leastcon$Latitude), col="black", cex=1, pch=16)
legend("topright", legend=c("Threatened", "Least Concern"), pch=c(4, 16), col=c("black", "black"))
dev.copy2pdf(file="NFLrastermap_status.pdf", width = 7, height = 5)

plot(rast$bio1, xlim=c(-65,-50), ylim=c(45,58), 1, main="", xaxt="n", yaxt="n", legend=F) #plotting the newfoundland populations to get an idea of where they are  
points(cbind(alldat$Longitude, alldat$Latitude), col=alldat$HWE.values2 == 0, cex=1, pch=20)
points(cbind(alldat$Longitude, alldat$Latitude), col=alldat$HWE.values2 == 1, cex=1, pch=17)

head(alldat)
```

