---
title: "ClimateRegressions"
author: "Katie Birchard"
date: '2018-04-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Now that we have the trends in climatic data, we can perform linear regressions on these trends with indicators of genetic variation in the salmon populations. First, I need to put everything together in one dataset:

```{r}
bigclim <- read.csv("ClimTrendData.csv") #downloading the climate dataset I made
View(bigclim)
bigclim2 <- data.frame(bigclim$code, bigclim$pastTemp, bigclim$CurTemp, bigclim$FutTemp) #taking only the columns I need
bigclim3 <- bigclim2[order(bigclim2$bigclim.code), ] #reordering it by alphabetizing the populations codes (this is so that the data easily aligns to the HWE dataset)

HWEvalues <- read.csv("Extracted_HWE.csv") #downloading the HWE dataset that contains values for allelic richness, expected heterozygosity, and HWE p-values
View(HWEvalues)
HWEvalues2 <- HWEvalues[order(HWEvalues$Populations), ] #reordering the populations by alphabetizing them so they line up with the climate dataset

HWEvalues2$pastTemp <- bigclim3$bigclim.pastTemp #adding past temp data
HWEvalues2$CurTemp <- bigclim3$bigclim.CurTemp #adding current temp data
HWEvalues2$FutTemp <- bigclim3$bigclim.FutTemp #adding future temp data

alldat <- HWEvalues2 #renaming the dataset to something more appropriate


#need climate trends as well (i.e. slopes of climate data over time)
trenddat <- data.frame(1:nrow(alldat), alldat$pastTemp, alldat$CurTemp, alldat$FutTemp)


#making a function for slope
slope  <-  function(x){
    if(all(is.na(x)))
        return(NA) #returns NA if missing data
    else
        return(coef(lm(I(1:3)~x))[2]) #uses a linear model to find the slope of 3 consecutive values (Past temp, Cur temp, and fut temp)
}

# now apply the function to all rows in trenddat and insert the slope into a new column called "trends"
trenddat$trends <- apply(trenddat[ ,c("alldat.pastTemp", "alldat.CurTemp", "alldat.FutTemp")], 1, slope)

alldat$ClimTrend <- trenddat$trends #adding the trend column to the big dataset

head(alldat) #yay, it worked!
View(alldat)
```

Now, we will see if there is a correlation between climate trends and allelic richness.

* We want the climate trends to be our predictor variable and allelic richness to be our outcome variable

* Basically trying to see if allelic richness can be "predicted" by changes in mean annual temperature

```{r}
# first, we will do a quick visualization with a scatterplot to see if there are any apparent trends:
scatter.smooth(x=alldat$ClimTrend, y=alldat$Allelic.Richness, main="Allelic Richness vs. Temp Trends")
# doesn't look like there is anything going on...

# Now I will quickly check for outliers in the data using a boxplot:
boxplot(alldat$ClimTrend, main="Temp Trends", sub=paste("Outlier rows: ", boxplot.stats(alldat$ClimTrend)$out))
#ClimTrend doesn't seem to have any outliers
boxplot(alldat$Allelic.Richness, main="Allelic Richness", sub=paste("Outlier rows: ", boxplot.stats(alldat$Allelic.Richness)$out))
#Allelic Richness has a few outliers...

boxplot.stats(alldat$Allelic.Richness)$out #6.41, 5.50, 4.85, 4.93, 6.22, 6.44 are all listed as outliers

#Now let's check for normality by doing a density plot
install.packages("e1071") #packages that measures skewedness in density plots
library(e1071)
plot(density(alldat$ClimTrend), main= "Density Plot: Temp Trends", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$ClimTrend), 2)))
# the plot for temp trends looks slightly bimodal...
# skewedness = -0.29

plot(density(alldat$Allelic.Richness), main= "Density Plot: Temp Trends", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$Allelic.Richness), 2)))
# wow, this is skewed by those outliers
# Skewedness = -1.7

#get a look at the histogram
hist(alldat$Allelic.Richness)

# Now lets look at the level of correlation between Allelic Richness and Temp Trends
## first with the outliers
cor(alldat$ClimTrend, alldat$Allelic.Richness) # R = -0.008830456

## now without the outliers
ARoutliers <- boxplot.stats(alldat$Allelic.Richness)$out
min(ARoutliers)
max(ARoutliers)

ARdrop <- alldat[!(alldat$Allelic.Richness <= max(ARoutliers)),]
cor(ARdrop$ClimTrend, ARdrop$Allelic.Richness) # R= -0.09794524
# there's a bit more of a negative correlation, but still not much

#how does this hist look:
hist(ARdrop$Allelic.Richness)
shapiro.test(ARdrop$Allelic.Richness) #data do not fit normality test (p=0.0003), skewed but can still do a least squares regression
shapiro.test(log10(ARdrop$Allelic.Richness)) #data not normal (p=2.89e=05)
shapiro.test(1/(ARdrop$Allelic.Richness)) #not normal (p=2.1e-06)


# lets see what the scatterplot looks like for this:
scatter.smooth(x=ARdrop$ClimTrend, y=ARdrop$Allelic.Richness, main="Allelic Richness vs. Temp Trends")
# yeahhhhhh, still not really looking like a correlation

# Now I'm making a linear model to get a better quanititative idea
lmAR <- lm(Allelic.Richness ~ ClimTrend, data=alldat)
print(lmAR) # gives intercept and slope
summary(lmAR) # gives summary statistics, p=0.914 (not significant), R^2=7.798e=05
resAR <- residuals(lmAR) #finding residuals
plot(alldat$ClimTrend, resAR) #plotting residuals(difference b/w observed value and predicted value)

#what about a linear model for the dataset with the outliers removed:
lmARdrop <- lm(Allelic.Richness ~ ClimTrend, data=ARdrop)
summary(lmARdrop) #p=0.2379, R^2=0.009593

#I want to make a ggplot figure to show the linear regression (or lack of)
library(ggplot2)

#customizing a ggplot theme

theme_katie<-function(base_size = 18, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace% 
    theme(axis.text = element_text(size = rel(0.8),colour="black"),
          axis.ticks = element_line(colour = "black"),
    			legend.key = element_rect(fill="transparent", colour=NA),
    			panel.background = element_blank(),
    			legend.background = element_rect(fill="transparent", colour=NA),
          legend.position = "none",
          panel.border = element_rect(fill = NA, colour = NA), 
          panel.grid.major = element_line(colour = NA,size = 0), 
          panel.grid.minor = element_line(colour = NA,size = 0), 
          axis.line = element_line(colour ="black")
    )
}

install.packages("ggpmisc") #installing package that will allow me to put the p-values and R^2 values directly on the plots
library(ggpmisc)

ARbigLM <- ggplot(ARdrop, aes(y=Allelic.Richness, x=ClimTrend)) +
	geom_point(colour="cyan3", size=5, alpha=0.6) +
	#geom_smooth(method=lm, colour="cyan4") +
	theme_katie() +
	ylab("Allelic Richness") +
	xlab("Trend in Annual Temperature over Time") +
	stat_fit_glance(method="lm", geom="text", aes(label=paste0("p = ", round(..p.value.., 3))), label.x.npc = 0.8, label.y.npc = 0.2) +
	stat_poly_eq(formula = ARdrop$Allelic.Richness ~ ARdrop$ClimTrend, 
                aes(label = paste0(..rr.label..)), label.x.npc = 0.8, label.y.npc = 0.1, parse = TRUE)
print(ARbigLM)
ggsave("AR_BIG.pdf")
```

```{r}

#what if we log transform it?
lmARlog <- lm(log10(Allelic.Richness) ~ log10(ClimTrend), data=alldat)
print(lmARlog)
summary(lmARlog) #p=0.0853

ARlogLM <- ggplot(alldat, aes(y=log10(Allelic.Richness), x=log10(ClimTrend))) +
	geom_point(colour="cyan3", size=5, alpha=0.6) +
	geom_smooth(method=lm, colour="cyan4") +
	theme_katie() +
	ylab("Log of Allelic Richness") +
	xlab("Trend in Annual Temperature over Time") +
	stat_fit_glance(method="lm", geom="text", aes(label=paste0("p = ", round(..p.value.., 3))), label.x.npc = 0.8, label.y.npc = 0.2) +
	stat_poly_eq(formula = log10(alldat$Allelic.Richness) ~ log10(alldat$ClimTrend), 
                aes(label = paste0(..rr.label..)), label.x.npc = 0.8, label.y.npc = 0.1, parse = TRUE)
print(ARlogLM)
ggsave("AR_BIG_log.pdf")
```

```{r}

#boostrap simulation b/w populations with neg climate trend vs. pos clim trend?
## trying to find out if mean allelic richness is significantly different between pos and neg climate trends

# make new column in alldat for categorical variable - clim pos/neg
alldat$ClimSlope <- NA
head(alldat)

#Making a loop to assign either a 0 or a 1 to the populations for my new column
for (i in 1:nrow(alldat)) {
	if (alldat$ClimTrend[i] > 0) {
		alldat$ClimSlope[i] <- "Pos"
	} 
	if (alldat$ClimTrend[i] < 0) {
		alldat$ClimSlope[i] <- "Neg"
	} 
}

Iters <- 1000 #we want to take a random sample 1000 times
BootOut <- rep(NA, Iters) #making a column with 1000 levels for the output of my bootstrap analysis
ClimPos <- alldat$Allelic.Richness[grep("Pos", alldat$ClimSlope)] #taking all the allelic richness values that have positive climate slope values
ClimNeg <- alldat$Allelic.Richness[grep("Neg", alldat$ClimSlope)] #taking all the allelic richness values that have a negative slope
for (i in 1:Iters){
	BootOut[i] <-
		mean(sample(ClimPos, length(ClimPos), replace=T)) - #taking mean of a sample of the allelic richness values for positive slopes
		mean(sample(ClimNeg, length(ClimNeg), replace=T)) #subtracting the mean of a sample of allelic richness values for negative slopes
}
head(BootOut) 


qplot(BootOut) # look at the distribution of values

XBootOut <- mean(BootOut) #mean of distribution is -0.1725703
CIlow <- sort(BootOut)[floor(length(BootOut)*0.025)]
CIhigh <- sort(BootOut)[ceiling(length(BootOut)*0.975)]
CIboot <- round(c(CIlow, CIhigh),2)

#To find p-value, need the mean difference of the raw data:
RawDat <- aggregate(alldat$Allelic.Richness, list(alldat$ClimSlope), mean)
RawDat

RawDiff <- RawDat$x[2] - RawDat$x[1]
qplot(BootOut) + theme_bw() + geom_vline(aes(xintercept=RawDiff), colour="blue")

p <- sum(BootOut>RawDiff)/length(BootOut)
paste("The estimated p-value is:", p) #p=0.512, so not really a big difference here


```

Next, we will see if there is a correlation between climate trends and expected heterozygosity:

```{r}
# first, we will do a quick visualization with a scatterplot to see if there are any apparent trends:
scatter.smooth(x=alldat$ClimTrend, y=alldat$Expected.Het, main="Expected Heterozygosity vs. Temp Trends")
# looks like expected heterozygosity is pretty stable with temperature fluctuations...

# Now I will quickly check for outliers in the data using a boxplot:
boxplot(alldat$Expected.Het, main="Expected Heterozygosity", sub=paste("Outlier rows: ", boxplot.stats(alldat$Expected.Het)$out))
#There are quite a few outliers for expected heterozygosity...

boxplot.stats(alldat$Expected.Het)$out #0.78 0.79 0.77 0.70 0.77 0.69 0.66 0.78 0.79 0.79 0.79 0.77 are all listed as outliers

#Now let's check for normality by doing a density plot
plot(density(alldat$Expected.Het), main= "Density Plot: Expected Heterozygosity", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$Expected.Het), 2)))
# this is even more skewed than the allelic richness density plot
# Skewedness = -2.54

# Now lets look at the level of correlation between Expected Heterozygosity and Temp Trends
## first with the outliers
cor(alldat$ClimTrend, alldat$Expected.Het) # R = 0.1223153

## now without the outliers
HEoutliers <- boxplot.stats(alldat$Expected.Het)$out
min(HEoutliers)
max(HEoutliers)

HEdrop <- alldat[!(alldat$Expected.Het <= max(HEoutliers)),]
cor(HEdrop$ClimTrend, HEdrop$Expected.Het) # R= 0.09128052 --> correlation coefficient actually went down

# lets see what the scatterplot looks like for this:
scatter.smooth(x=HEdrop$ClimTrend, y=HEdrop$Expected.Het, main="Expected Heterozygosity vs. Temp Trends")
# no correlation here

# Now I'm making a linear model to get a better quanititative idea
lmHE <- lm(Expected.Het ~ ClimTrend, data=alldat)
summary(lmHE) # gives summary statistics, p=0.132, R^2=0.01496

#what if we log transform it?
lmHElog <- lm(log10(Expected.Het) ~ log10(ClimTrend), data=alldat)
summary(lmHElog) #p=0.007319, R^2=0.06714- SIGNIFICANT!!

#linear model without the outliers:
lmHEdrop <- lm(Expected.Het ~ ClimTrend, data=HEdrop)
summary(lmHEdrop) #p=0.2817, R^2=0.008332

HEbigLM <- ggplot(HEdrop, aes(x=ClimTrend, y=Expected.Het)) +
	geom_point(colour="cyan3", size=5, alpha=0.6) +
	#geom_smooth(method=lm, colour="cyan4") +
	theme_katie() +
	ylab("Expected Heterozygosity") +
	xlab("Trend in Annual Temperature over Time") + 
	stat_fit_glance(method="lm", geom="text", aes(label=paste0("p = ", round(..p.value.., 3))), label.x.npc = 0.8, label.y.npc = 0.2) +
	stat_poly_eq(formula = HEdrop$Expected.Het ~ HEdrop$ClimTrend, 
                aes(label = paste0(..rr.label..)), label.x.npc = 0.8, label.y.npc = 0.1, parse = TRUE)
print(HEbigLM)
ggsave("HE_BIG.pdf")
```

```{r}
# Now plot the log of HE
HElogLM <- ggplot(alldat, aes(y=log10(Expected.Het), x=log10(ClimTrend))) +
	geom_point(colour="cyan3", size=5, alpha=0.6) +
	geom_smooth(method=lm, colour="cyan4") +
	theme_katie() +
	ylab("Log of Expected Heterozygosity") +
	xlab("Trend in Annual Temperature over Time") +
	stat_fit_glance(method="lm", geom="text", aes(label=paste0("p = ", round(..p.value.., 3))), label.x.npc = 0.8, label.y.npc = 0.2) +
	stat_poly_eq(formula = log10(alldat$Expected.Het) ~ log10(alldat$ClimTrend), 
                aes(label = paste0(..rr.label..)), label.x.npc = 0.8, label.y.npc = 0.1, parse = TRUE)
print(HElogLM)
ggsave("HE_BIG_log.pdf")
```


Maybe there is an interaction going on between multiple values. I'm going to run multiple regressions now:
```{r}
#We want a dataset free of the outliers from both expected heterozygosity and allelic richness:
MultDrop <- ARdrop[!(ARdrop$Expected.Het <= max(HEoutliers)),]

#run a least squares regression:
lmMultDrop <- lm(Expected.Het*Allelic.Richness ~ ClimTrend, data=MultDrop)
summary(lmMultDrop) #p=0.653

#this analysis actually doesn't make much sense because allelic richness and He are related
```


Then, we will see if there is a correlation between climate trends and deviations from HWE:

```{r}
# first, we will do a quick visualization with a scatterplot to see if there are any apparent trends:
scatter.smooth(x=alldat$ClimTrend, y=alldat$HWE.Values, main="Deviations from HWE vs. Temp Trends")
# looks like there is no definitive relationship between temp trends and deviations from HWE...

# Now I will quickly check for outliers in the data using a boxplot:
boxplot(alldat$HWE.Values, main="Deviations from HWE", sub=paste("Outlier rows: ", boxplot.stats(alldat$HWE.Values)$out))
#no outliers...

range(alldat$HWE.Values)
hist(alldat$HWE.Values) #looks like there are 2 groups of individuals, one that deviates from HWE, and one that does not - should do logistical regression for this

# I'm going to make a nice histogram for the HWE p-values:
HWE_hist <- ggplot(data=alldat, aes(HWE.Values)) + 
	geom_histogram(breaks=seq(0, 1, by=0.05),
								 col="cyan4", #fill="cyan3"
								 aes(fill=..count..),
								 alpha=0.8) +
  scale_fill_gradient("Populations", low = "cyan1", high = "cyan3") +
	theme_katie() +
	labs(title = "Atlantic Canada") +
	labs(x= "Hardy Weinberg Equilibrium P-values", y="Number of Populations") #+
	#geom_density(col="cyan4")
print(HWE_hist)
ggsave("HWEhist_BIG.pdf")
```

```{r}

#Now let's check for normality by doing a density plot
plot(density(alldat$HWE.Values), main= "Density Plot: Deviations from HWE", ylab="Frequency", 
		 sub=paste("Skewness:", round(e1071::skewness(alldat$HWE.values), 2)))
# bimodal distribution

# I can make the HWE p values truly dichotomous by assigning p < 0.05 == 0, and p > 0.05 == 1 - then I can run a logistic regression on it to see if there is a correlation with temperature trends

#I'm making a new row to put the binary HWE values in:
alldat$HWE.values2 <- NA
head(alldat)

#Making a loop to assign either a 0 or a 1 to the populations for my new column
for (i in 1:nrow(alldat)) {
	if (alldat$HWE.Values[i] >= 0.05) {
		alldat$HWE.values2[i] <- 1
	} else {
		alldat$HWE.values2[i] <- 0
	} 
}

head(alldat) #my new column is populated, so now I can perform a logistic regression

mod1<-glm(HWE.values2~ClimTrend, data=alldat, family=binomial(link="logit"))
summary(mod1) # p=0.0876, not significant
# endangered populations tend to be further south, so maybe we should try with latitude

# taking the latitude and longitudes out of the bigclim dataset and putting them into alldat:
#first I have to reorder bigclim by population so it's in same order as alldat:
bigclim[order(bigclim$code), ]
alldat$Latitude <- bigclim$Latitude
alldat$Longitude <- bigclim$Longitude
head(alldat)

# Running a logistic regression with HWE deviations and latitude
mod2<-glm(HWE.values2~Latitude, data=alldat, family=binomial(link="logit"))
summary(mod2)

#Latitude + longitude
mod3 <- glm(HWE.values2~Latitude+Longitude, data=alldat, family=binomial(link="logit"))
summary(mod3)

#just longitude
mod4 <- glm(HWE.values2~Longitude, data=alldat, family=binomial(link="logit"))
summary(mod4)

# What about if there is a relationship b/w deviations from HWE and endangerment status?
StatusDat <- read.csv("EndangermentStatusNFL.csv")
head(StatusDat)

chisq.test(StatusDat$Status, alldat$HWE.values2)
```

As with the Newfoundland dataset, these Hardy-Weinberg p-values also appear to be bimodally distributed - I'm going to plot these on a raster map

```{r}
#plot on map 

library(raster)

rast <- getData("worldclim", var="bio", res=2.5) #making raster object, using biological variables, lowest resolution
rast <- rast[[c(1)]] #taking out only the mean annual temperature variable (taken b/w 1950-2000)

head(bigclim) #not in same order as alldat
bigclim4 <- bigclim[order(bigclim$code), ]
head(bigclim4) #now in the same order as alldat

realcoords <- read.csv("mec12972-sup-0001-TableS1-S5.csv") #other dataset only has rounded lats/longs
head(realcoords)
realcoords2 <- realcoords[order(realcoords$code), ]
head(realcoords2) #now in same order as alldat

alldat$Latitude <- realcoords2$Latitude
alldat$Longitude <- realcoords2$Longitude

coords <- data.frame(x=alldat$Longitude, y=alldat$Latitude) #making dataframe for my lats and longs

points <- SpatialPoints(coords, proj4string = rast@crs) #determining the datapoints by specifying my lat and long

values <- extract(rast, points) #extracting the mean annual temperature for those lats and longs

df <- cbind.data.frame(coordinates(points), values) #putting lat, long, and mean annual temp into a dataframe

head(df) #this dataframe includes temperatures for all specific lats and longs for all populations

projection(rast)=CRS("+init=epsg:4326") #want to make sure I'm using the right coordinate system

deviate <- alldat[which(alldat$HWE.values2 == 0), ]
nrow(deviate)
nodeviate <- alldat[which(alldat$HWE.values2 == 1), ]
nrow(nodeviate)

range(alldat$Longitude) #-71.73 to -53.35
range(alldat$Latitude) #43.86 to 58.82

rastNFL <- plot(rast$bio1, xlim=c(-75,-50), ylim=c(40,60), 1, main="", xaxt="n", yaxt="n", legend=F, box=F, axes=F) #plotting the newfoundland populations to get an idea of where they are 
rastNFL
title("Atlantic Canada Populations", adj=0.75)
points(cbind(deviate$Longitude, deviate$Latitude), col="black", cex=1, pch=4) #adding the points
points(cbind(nodeviate$Longitude, nodeviate$Latitude), col="black", cex=1, pch=16)
legend("topright", legend=c("Deviation", "No Deviation"), pch=c(4, 16), col=c("black", "black"))
dev.copy2pdf(file="NFLrastermap_dev.pdf")
```

