---
title: "PophelperBigFile"
author: "Lila Colston-Nepali"
date: '2018-04-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Pophelper allows you to summarize your STRUCTURE results- find the best K value, and plot the data

For mac users: you will need to install https://www.xquartz.org/ in order to use the package Cairo

```{r}
# install dependencies and devtools
install.packages(c("Cairo","ggplot2","gridExtra","gtable","tidyr","devtools"),dependencies=T)

library(c(devtools, ggplot2, gridExtra, gtable, tidyr, devtools, Cairo))

# install pophelper package from GitHub
devtools::install_github('royfrancis/pophelper')

# load library for use
library(pophelper)


```

The readQ function makes a 'qlist'- a list of the Q-matrices, or the probabilities of 
individuals to be assigned to a genetic cluster

This qlist can then be used in further steps that will allow us to calculate the 'best K'-
  the number of genetic clusters that best fits the data set
  
```{r}

#Make an object that contains the relevant STRUCTURE output files
BigFileStrOut <- readQStructure("BIGFILEoutput/")


#The next steps prepare the data to analyze best K
tabulates <- tabulateQ(qlist = BigFileStrOut)
summarized <- summariseQ(data = tabulates)

# The Evanno method is a commonly accepted method to determine best K, this function calculates it
evannoMethodStructure(summarized)


```

The best K value was 13- as the Delta K value for K =13 was the highest.

Now that we know the best K value, we can plot this output!
The plotQ() function is the basic function for plotting
This function is super cool though- as with ggplot2, there are so many ways to customize
your plots!!

```{r}
#plotting

BESTK <- readQStructure("STRUCOUT/bigfilestroutput13_f")

labelsBF <- read.delim("BigFileLabels.txt",header=FALSE,stringsAsFactors=FALSE)

plotQ(BESTK, showsp=F, clustercol = c("aquamarine4", "darkslategrey", "darkcyan", "cyan3", "medium aquamarine", "lightseagreen", "paleturquoise4", "medium turquoise", "medium seagreen", "seagreen4", "springgreen4", "turquoise4", "palegreen4"),
      showlegend = T, legendkeysize = 6, showyaxis = T, 
      grplab = labelsBF, grplabsize = 1, linesize = 0.3, pointsize = 1, showdiv = FALSE,
      grplabspacer = -0.2, grplabheight = 1, linetype = 1, font = "Arial", ordergrp = T)


```

Plotting details:
There are SO many ways to customize your plot. Here is the details of some of my customizations.
clustercol defines the colours

We want the colours to be consistent amongst our plots so we coordinated this.

You can have divider lines between groups, but I find it looks best without.
The labels (see below) can illustrate this fine.

ordergrp is REALLY useful! It allows you to sort your individuals the way you want- I set it based on our labels.

showsp lets you decide whether you want the strip panel on the side (that says K = N). 
Aesthetically, I thought it looked best without it, so I set it to FALSE.

showlegend displays the legend showing the two clusters.You can change the size of the legend.

showyaxis shows the scale from 0 to 1 of posterior probability membership to each cluster.
I chose to include it.

grplab allows you to label the plot by group (in our case region), for better visualization.
You can customize these labels with size, space and height. I played around with the parameters
until I found ones that I thought looked good.
You can also change the size of the label lines etc.
