---
title: "NFLmap for Endangered"
author: "Katie Birchard"
date: '2018-04-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We want to know which populations are threatened and which ones aren't for our analyses. There are no specified coordinates online for this, but there is a map of COSEWIC endangerment status online @ http://www.asf.ca/endangered-populations.html. I'm going to make a map to compare our population locations to the locations of threatened species in Newfoundland.

```{r}
NFLclim <- read.csv("ClimTrendDataNFL4.csv", header=T)
View(NFLclim)
NFLclim1 <- data.frame(NFLclim$River, NFLclim$Latitude, NFLclim$Longitude)
names(NFLclim1) <- c("River", "Latitude", "Longitude")
head(NFLclim1)
# NFLclim <- NFLclim[,apply(NFLclim, 2, function(x) { sum(!is.na(x)) > 0 })] # function to remove columns that just contain NA
BIGclim <- read.csv("mec12972-sup-0001-TableS1-S5.csv")
View(BIGclim)
BIGclim1 <- data.frame(BIGclim$Name, BIGclim$Latitude, BIGclim$Longitude)
names(BIGclim1) <- c("River", "Latitude", "Longitude")
head(BIGclim1)

data <- rbind(NFLclim1,BIGclim1)
View(data)
data1 <- data[!duplicated(data), ] #getting rid of duplicate rivers
nrow(data1)
data1$code <- 1:nrow(data1)

data2 <- data1[order(data1$River), ]
View(data2)

#making a map of the points that is not a raster map
range(NFLclim$Latitude)
range(NFLclim$Longitude)

minlat <- min(NFLclim$Latitude) 
maxlat <- max(NFLclim$Latitude)  
minlong <- min(NFLclim$Longitude) - 2
minlong
maxlong <- max(NFLclim$Longitude) + 2
maxlong
View(NFLclim)

is.numeric(minlat)

library(maps)
library(ggplot2)
world_map <- map_data("world")
#tried to subset the world map all in one go, but it didn't like this, so had to do 2 steps
subsetMap <- subset(world_map, #world_map$long > minlong & 
											world_map$long > maxlong &
											world_map$lat > minlat) #&
											#world_map&lat < maxlong) 

subsetMap2 <- subset(subsetMap, world_map$long < minlong & world_map$lat < maxlat)
subsetMap2

baseplot <- ggplot() + coord_fixed() +
  xlab("") + ylab("")

baseworld_rough <- baseplot + #making a base world plot, and specifying my coordinates
	geom_polygon(data=world_map, aes(x=long, y=lat, group=group),
																					colour="green", fill="light green") +
	coord_quickmap(xlim = c(minlong,maxlong), ylim=c(minlat,maxlat))

baseworld_rough

cleanup <- theme(panel.grid.major = element_blank(), #making it pretty
				panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = 'lightblue1', colour = 'lightblue1'), 
        axis.line = element_line(colour = "white"), legend.position="none",
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())

base_world <- baseworld_rough + cleanup #adding the prettiness to my rough map
base_world

NFLclim$mapcode <- c(1:nrow(NFLclim)) #coding the populations so it isn't a huge mess when I add the labels to the map

library(ggrepel) #need this package to add labels
NFLmap_coloured <- base_world + #adding the populations to my map
  geom_point(data=NFLclim,aes(x=Longitude, y=Latitude, colour=River), size=2, alpha=I(0.7)) +
	geom_text_repel(data = NFLclim, aes(Longitude, Latitude, label = mapcode), size = 3,
                  box.padding = unit(0.1, 'lines'), force = 0.5)

NFLmap_coloured
dev.copy2pdf(file="NFLmap_coloured.pdf", width = 7, height = 5) #saving the map to a pdf
```

Now I can compare the two maps to see where the populations we have lie on the COSEWIC endangerment map.
Using this resource, plus supplementary resources for the ones that aren't clear on the map, we will make another column in the dataset for threatened/non-threatened.

```{r}
NFLclim$Status <- NA
View(NFLclim)
NFLclim$Code <- c("NL-ALX", "NL-BBR", "BDN", "NL-BVB", "NL-BIG", "BBA", "NL-BER", "BSB", "NL-BDB", "NL-CMP", "CCR", "CNR", "CMB", "NL-DWB", "DHB", "NL-ENG", "NL-RBR", "NL-FLB", "GBI", "GAR", "GBB", "NL-COD", "GNR", "NL-GRB", "GRR", "NL-HUR", "NL-IAB", "NL-IBB", "IAM", "NL-JUB", "NL-KEN", "LPR", "NL-LJB", "LRD", "LSR", "NL-LOM", "LHR", "NL-MNR", "NL-MBA", "NHR", "NEB", "NPR", "NWT", "OBB", "NL-PGB", "NL-PIN", "NL-PHR", "NL-RWR", "NL-ROP", "RKR", "NL-ROF", "NL-SRN", "SLR", "NL-SAN", "FBS", "NL-GSB", "SPR", "SWB", "SBM", "NL-SGR", "NL-SMR", "STS", "STN", "NL-TYB", "NL-TNR", "NL-WAB", "WBR")
View(NFLclim)
NFLendangered <- NFLclim[-grep("(\\w)(\\w)-", NFLclim$Code), ] #using regex to get the endangered populations only - these populations have a 3 letter code because they are all from one dataset (Bradbury et al. 2015)
View(NFLendangered)

#can make ifelse statement using regex to assign endangerment status:
for (i in 1:nrow(NFLclim)) {
	if (length(grep("(\\w)(\\w)-", NFLclim$Code[i])) == T) {
		NFLclim$Status[i] <- "Least Concern"
	} else {
		NFLclim$Status[i] <- "Threatened"
	}
}

#yay it worked!

write.csv(NFLclim, "EndangermentStatusNFL2.csv")
```

```{r}

NFLclim1 <- data.frame(NFLendangered$River, NFLendangered$Latitude, NFLendangered$Longitude)
names(NFLclim1) <- c("River", "Latitude", "Longitude")
head(NFLclim1)


minlat <- min(data2$Latitude) 
maxlat <- max(data2$Latitude)  
minlong <- min(data2$Longitude) - 2
minlong #-73.73
maxlong <- max(data2$Longitude) + 2
maxlong # -51.28
View(data2)

is.numeric(minlat)

NFL <- data2[which(data2$code <= 32), ]
nrow(NFL)
View(NFL)
BIG <- data2[which(data2$code > 32), ]
nrow(BIG)

map_coloured <- base_world + #adding the populations to my map
  geom_point(data=NFL,aes(x=Longitude, y=Latitude, colour="cyan3"), size=2, alpha=I(0.7)) +
	geom_point(data=BIG,aes(x=Longitude, y=Latitude, colour="darksalmon"), size=2, alpha=I(0.7))

map_coloured #not quite what I wanted

NFLclim1 <- data.frame(NFLendangered$River, NFLendangered$Latitude, NFLendangered$Longitude)
names(NFLclim1) <- c("River", "Latitude", "Longitude")
head(NFLclim1)

data_fix <- rbind(NFLclim1,BIGclim1)
View(data_fix)
data_fix$code <- 1:nrow(data_fix)

NFL <- data_fix[which(data_fix$code <= 31), ]
nrow(NFL)
View(NFL)
BIG <- data_fix[which(data_fix$code > 31), ]
nrow(BIG)

map_coloured2 <- base_world + #adding the populations to my map
  geom_point(data=NFL,aes(x=Longitude, y=Latitude, colour="cyan4"), size=2, alpha=I(0.7)) +
	geom_point(data=BIG,aes(x=Longitude, y=Latitude, colour="darksalmon1"), size=2, alpha=I(0.7)) +
	scale_colour_manual("", breaks = c("Newfoundland", "Atlantic Canada"), values = c("darksalmon", "cyan3"))

map_coloured2
dev.copy2pdf(file="intromap.pdf")

```

